A full approximation scheme multilevel method \\ for nonlinear variational inequalities

### 2023-07-19 paper/fascd.tex
eliminate Example 3.4, and tighten adjacent text

The basic CD iterations are meaningful whether or not we are in the optimization case $f=F'$, and when $f$ is nonlinear, non-local, or defined only on $\cK$.  However, practical and efficient FE implementation for such general operators $f$ seems not to have been addressed in the literature.  References \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem, wherein $f=F'$ is linear, local, and defined on all of $\mathcal{V}$. Substantial implementation simplifications are available in such cases.

\begin{example}  \label{ex:fnice} Suppose $f:\cV \to \cV'$ is linear, and suppose it is local in the sense that for any $z\in\mathcal{V}$ the value $\ip{f(\phi)}{z}$ can be computed by an integral over the support of $\phi$.  Considering \pr{cd-add} for concreteness, by linearity the VI can be rewritten
\begin{equation}
\ip{f(e_i)}{v_i-\hat w_i} \ge \ip{\tilde\ell}{v_i-\hat w_i}, \label{eq:linearlocalvi}
\end{equation}
for all $v_i \in \mathcal{K}_i$, where $\tilde\ell = \ell - f(u)$ incorporates the initial residual $f(u)$.  Note $e_i = \hat w_i - \Pi_i u \in \cV_i$ is not generally admissible---generally $e_i \notin \cK$---but ``$f(e_i)$'' in \eqref{eq:linearlocalvi} is meaningful because $f$ is defined on all of $\cV$.
\end{example}

\pef{I confess I just don't get the point of this example. Why is this a substantial implementation simplification? I am sorry I am being dumb. Maybe you can explain it to me on a video call and I will write it in a way that I would have understood it?}
\elb{Not being dumb.  I am describing an application I am headed toward.  FAScd applies to both differential \emph{and integral} operators.  (We don't try it out in the latter case, but it is a feature that it does not care about the locality of your operator, unlike monotone multigrid.)  It applies to operators where $f(e_i)$ makes no sense because $e_i\notin \mathcal{K}$ (SIA is an example IMO).  And it does FAS coarse corrections because we don't assume linearity of $f$.}

When $f$ has the locality of Example \ref{ex:fnice}, an FE approximation of VI \eqref{eq:vi} can be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over basis function supports.\footnote{Compare equations (2.31) and (2.32) in \cite{Farrelletal2021}, which describe the same sense of locality.}  However, for general problems lacking the convenience of Example \ref{ex:fnice}, certain additional ideas are needed for practical implementation of CD-type iterations.  Section \ref{sec:vcycle} constructs such an iteration for a general operator $f:\cK\to\cV'$ by applying the full approximation scheme (FAS) idea of Brandt \cite{Brandt1977}.

From now on we restrict to the case of \emph{box constraints} \cite{BensonMunson2006,FerrisPang1997}, equivalently \emph{bi-obstacle problems}.  Suppose $\Omega \subset \RR^d$ is open and let $\mathcal{V}=W^{1,p}(\Omega)$.  The \emph{lower obstacle} $\underline{\gamma}$ and \emph{upper obstacle} $\overline{\gamma}$ are given measurable functions, defined on the closure $\overline{\Omega}$, with values in the extended real line $\tilde\RR = [-\infty,+\infty]$, satisfying the following properties: $\underline{\gamma} : \overline{\Omega} \to [-\infty,+\infty)$, $\overline{\gamma} : \overline{\Omega} \to (-\infty,+\infty]$, and $\underline{\gamma} \le \overline{\gamma}$ on $\overline{\Omega}$.  Also suppose $\partial\Omega$ is split into disjoint, measurable sets, $\partial\Omega = \partial_D \Omega \cup \partial_N \Omega$, with $\partial_D \Omega$ of positive Hausdorff measure, and $\partial_N \Omega$ possibly empty.  Dirichlet data\footnote{Dirichlet boundary values are in the trace sense \cite{Evans2010}, and $g_D$ is assumed to be as regular as needed.} is given by $g_D:\partial_D \Omega \to \RR$, and we assume $\underline{\gamma} \le g_D \le \overline{\gamma}$ on $\partial_D \Omega$.  Now define
\begin{equation}
\cK = \left\{w\,:\,\underline{\gamma} \le w \le \overline{\gamma} \, \text{ and }\, w\big|_{\partial_D \Omega} = g_D\right\} \subset \cV =W^{1,p}(\Omega), \label{eq:originalconstraintset}
\end{equation}
the closed and convex \emph{constraint set}.  Note that we are treating inequality constraints and strongly-imposed Dirichlet boundary conditions in a unified manner.  Any Neumann conditions over $\partial_N \Omega$ are imposed weakly in the definition of the operator $f$, i.e.~by an integral over $\partial_N\Omega$.

In the next Sections we numerically solve continuum VI problem \eqref{eq:vi} over a constraint set of the form \eqref{eq:originalconstraintset}.  In fact, given continuous $f:\cK \to \cV'$, and $\ell \in \cV'$, from now on we assume that the following is well-posed for $u^\star\in \cK$:
\begin{equation}
\ip{f(u^\star)}{v-u^\star} \ge \ip{\ell}{v-u^\star} \qquad \text{for all } v\in \cK. \label{eq:boxdirichletvi}
\end{equation}
Note that $v-u^\star \in W_0^{1,p}(\Omega)$.

>>>>

The basic CD iterations are meaningful whether or not we are in the optimization case $f=F'$, and even if $f$ is nonlinear, non-local, or defined only on $\cK$.  However, practical and efficient FE implementation for general operators $f$ seems not to have been addressed in the literature.  References \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem, wherein $f=F'$ is linear, local, and defined on all of $\mathcal{V}$.  In Section \ref{sec:vcycle} we will extend CD-type iterations to nonlinear operators by applying the full approximation scheme (FAS) idea of Brandt \cite{Brandt1977}, and then illustrate its application to non-optimization operators in Examples \ref{ex:results:advdiff} and \ref{ex:results:sia}.  In the latter Example $f$ is defined only over a subset $\cK$.

From now on we restrict to the case of \emph{box constraints} \cite{BensonMunson2006,FerrisPang1997}, equivalently \emph{bilateral obstacle problems}.  Suppose $\Omega \subset \RR^d$ is open and let $\mathcal{V}=W^{1,p}(\Omega)$.  The \emph{lower obstacle} $\underline{\gamma}$ and \emph{upper obstacle} $\overline{\gamma}$ are given measurable functions, defined on the closure $\overline{\Omega}$, with values in the extended real line $\tilde\RR = [-\infty,+\infty]$, satisfying the following properties: $\underline{\gamma} : \overline{\Omega} \to [-\infty,+\infty)$, $\overline{\gamma} : \overline{\Omega} \to (-\infty,+\infty]$, and $\underline{\gamma} \le \overline{\gamma}$ on $\overline{\Omega}$.  Also suppose $\partial\Omega$ is split into disjoint, measurable sets, $\partial\Omega = \partial_D \Omega \cup \partial_N \Omega$, with $\partial_D \Omega$ of positive Hausdorff measure, and $\partial_N \Omega$ possibly empty.  Dirichlet data\footnote{Dirichlet boundary values are in the trace sense \cite{Evans2010}, and $g_D$ is assumed to be as regular as needed.} is given by $g_D:\partial_D \Omega \to \RR$, and we assume $\underline{\gamma} \le g_D \le \overline{\gamma}$ on $\partial_D \Omega$.  Now define
\begin{equation}
\cK = \left\{w\,:\,\underline{\gamma} \le w \le \overline{\gamma} \, \text{ and }\, w\big|_{\partial_D \Omega} = g_D\right\} \subset \cV =W^{1,p}(\Omega), \label{eq:originalconstraintset}
\end{equation}
the closed and convex \emph{constraint set}.  Note that we are treating inequality constraints and strongly-imposed Dirichlet boundary conditions in a unified manner.  (Any Neumann conditions over $\partial_N \Omega$ are imposed weakly by the operator $f$, via an integral over $\partial_N\Omega$.)  Next we will numerically solve continuum VI problem \eqref{eq:vi} over a constraint set \eqref{eq:originalconstraintset}.  In fact, given continuous $f:\cK \to \cV'$, and $\ell \in \cV'$, we will assume that the following is well-posed for $u^\star\in \cK$:
\begin{equation}
\ip{f(u^\star)}{v-u^\star} \ge \ip{\ell}{v-u^\star} \qquad \text{for all } v\in \cK. \label{eq:boxdirichletvi}
\end{equation}
Note that $v-u^\star \in W_0^{1,p}(\Omega)$.
